---
title: "Homework #5: Density Estimation" 
author: "**Your Name Here**"
date: "Due: Wed Mar 24 | 10:55am"
output: R6018::homework
---

**SYS 4582/6018 | Spring 2021 | University of Virginia **

*******************************************
```{r config, echo=FALSE}
source(system.file("config/hw_config.R", package="R6018")) # knitr settings
# options(dplyr.summarise.inform = FALSE)  # ignore dplyr message about grouping
```


# Required R packages and Directories

### {.solution}
```{r packages, message=FALSE, warning=FALSE}
data.dir = 'https://mdporter.github.io/SYS6018/data/' # data directory
library(R6018)     # functions for SYS-6018
library(tidyverse) # functions for data manipulation   
library(ks)        # functions for KDE
```



# Problem 1 Geographic Profiling

Geographic profiling, a method developed in criminology, can be used to estimate the [home location (roost) of animals](https://www.sciencedirect.com/science/article/pii/S0022519305004157) based on a collection of sightings. The approach requires an estimate of the distribution the animal will travel from their roost to forage for food. 

A sample of 283 distances that pipistrelle bats traveled (in meters) from their roost can be found at: 
<https://mdporter.github.io/SYS6018/data/geo_profile.csv>


One probability model for the distance these bats will travel is:
\begin{align*}
f(x; \theta) = \frac{x}{\theta} \exp \left( - \frac{x^2}{2 \theta} \right)
\end{align*}
where the parameter $\theta > 0$ controls how far they are willing to travel. 


## a. Derive the MLE for $\theta$ (i.e., show the math). 

### {.solution}

$$L(\theta)=\prod_{i=1}^nf(x_i,\theta)$$
$$log(L(\theta)) = \sum_i^n(\log{x_i} - \log{\theta} - \frac{x_i^2}{2\theta})$$
$$\frac{\partial{\log{L(\theta)}}}{\partial\theta} = \frac{n}{\theta} - \sum_i^n{\frac{x_i^2}{2*\theta^2}}=0$$
$$\frac{\sum_i^n{x_i^2}}{2*\theta^2} = \frac{n}{\theta}$$
$$\theta = \frac{\sum_i^nx_i^2}{2n}$$


## b. What is the MLE of $\theta$ for the bat data? (Use results from a, or use computational methods.) 

### {.solution}

```{r}

dataReader = read.csv("https://mdporter.github.io/SYS6018/data/geo_profile.csv",header = FALSE)

res.1b.sumvsq = sum(dataReader$V1*dataReader$V1)

res.1b = res.1b.sumvsq / (2*nrow(dataReader))
res.1b
```




## c. Using the MLE value of $\theta$ from part b, compute the estimated density at a set of evaluation points between 0 and 8 meters. Plot the estimated density.

### {.solution}

```{r}
func.1c <- function(x , theta){(x/theta)*exp(-x*x/(2*theta))}
x.seq = seq(0,8,length = 200)

dataplot.1c = tibble(x = x.seq,y=func.1c(x.seq,res.1b))

plot.1c.para = dataplot.1c %>% 
  ggplot(aes(x,y)) + geom_point() + geom_line()

plot.1c.para
```


## d. Estimate the density using KDE. Report the bandwidth you chose and produce a plot of the estimated density. 

### {.solution}

```{r}
res.kde.1d = kde(dataReader$V1)
res.kde.1d$h
plot(res.kde.1d,las = 1)
```




## e. Which model do you prefer, the parametric or KDE? 

### {.solution}

I prefer the KDE model, from the plot we can see, the parametric tend to have too large binary which tend to hide the detail features of the distribution. 
```{r}
ggplot(dataReader,aes(x=V1)) + geom_bar(color="Blue", fill="skyblue") + geom_density(color = "red")
```



# Problem 2: Interstate Crash Density

Interstate 64 (I-64) is a major east-west road that passes just south of Charlottesville. Where and when are the most dangerous places/times to be on I-64? The crash data (link below) gives the mile marker and fractional time-of-week for crashes that occurred on I-64 between mile marker 87 and 136 in 2016. The time-of-week data takes a numeric value of *\<dow\>.\<hour/24\>*, where the dow starts at 0 for Sunday (6 for Sat) and the decimal gives the time of day information. Thus `time=0.0417` corresponds to Sun at 1am and `time=6.5` corresponds to Sat at noon). 

- **Crash Data**: <`r file.path(data.dir, 'crashes16.csv')`>


## a. Extract the crashes and make a scatter plot with mile marker on x-axis and time on y-axis. 


### {.solution}

```{r message=FALSE, warning=FALSE}

dataReader2a = readr::read_csv("https://mdporter.github.io/SYS6018/data//crashes16.csv") 

dataReader2a %>% ggplot(aes(mile,time)) + geom_point()
```



## b. Use KDE to estimate the *mile marker* density. 

- Report the bandwidth. 
- Plot the density estimate.

### {.solution}

```{r}
res.kde.2b = kde(dataReader2a$mile)
res.kde.2b$h
plot(res.kde.2b,las = 1)
```



## c. Use KDE to estimate the temporal *time-of-week* density. 

- Report the bandwidth. 
- Plot the density estimate. 

### {.solution}

```{r}

res.kde.2c = kde(dataReader2a$time)
res.kde.2c$h

plot(res.kde.2c,las =1 )
```




## d. Use KDE to estimate the bivariate mile-time density. 

- Report the bandwidth parameters.
- Plot the bivariate density estimate. 

### {.solution}

```{r}
H1 = Hscv(dataReader2a)
res.2d.kde = kde(dataReader2a,H = H1)


plot(res.2d.kde,
          cont = c(5,10,20,30,40,50,60),                        # set contour levels
     # display = "filled.contour",                # use filled contour
     las=1, xlim = c(80, 140), ylim=c(-1, 8)
     )
points(dataReader2a, pch = 19,cex = 0.5, col='grey60')
grid()
```




## e. Based on the estimated density, approximate the most dangerous mile marker and time-of-week. 

### {.solution}

```{r}

plot(res.2d.kde,
          cont = seq(1,30,by=3),                        # set contour levels
     # display = "filled.contour",                # use filled contour
     las=1, xlim = c(80, 140), ylim=c(-1, 8)
     )
points(dataReader2a, pch = 19,cex = 0.5, col='grey60')
grid()
```
From the above plot we can see:

- The most dangerous miles marker and time-of-week is around $~(117,1.7)$. Mile marker : 117, on Monday 4.50pm
- second most dangerous miles marker and time-of week is around $~(103,5.5)$. Mile marker 103, on Friday Noon 
		